Supplementary Materials

### 2023-06-27 manuscript/supplementary.tex
ms updates

\documentclass[11pt]{article}
\RequirePackage{fullpage}
%\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
%\RequirePackage[osf]{mathpazo}
\let\temp\rmdefault
\RequirePackage{mathpazo}
\let\rmdefault\temp

@@

@@

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

@@

\section{Theory}
\subsection{Overview of Quantitative Genetic Models of Effective Population Size}

@@

The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
x_{0,i}$, though to indicate that we treat this as fixed, we use $p_0 = x_0$.
Then, the allele frequency change is,

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \\
  \Delta x_1 &= K_1 + H_1
\end{align}

@@

Now, let us look the variance of $\var(\Delta x_1)$ over evolutionary
replicates. Since the Mendelian segregation and the offspring random process
are independent,

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1).
\end{align}
%
Looking at each term,

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var\left(x_{0,i} (k_{0,i} - 1) \right)
\end{align}
%
where we ignore the covariance terms due to the sum, since these are on order
$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
the neutral alleles assort independently into diploids with respect to their
fitness, so we can simply the variance as

\begin{align}
  \var(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \var(x_{0,i}) \var(k_{0,i} - 1)  \\
            &\approx \frac{1}{4N} \var(x_{0,i}) \var(k_{0,i}).
\end{align}

Assuming no correlation between parental gametes (e.g. no inbreeding),
$\var(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\var(k) := \var(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\var(k) \approx 2$. Then, the
variance in allele frequency change due to non-heritable offspring number
variation is,

\begin{align}
  \var(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4}.
\end{align}

The Mendelian noise variance term can be derived similarly. First, the sum over
each parent's transmitted gametes can be simplified by noting that parents are
exchangeable over evolutionary replicates with respect to their contribution to
this term. In a constant population size, the double summation over parents and
their offspring can be replaced by a summation over offspring, since both sum
$N$ exchangeable terms. Then, note that $\var(\delta_{i,j}) = \nicefrac{1}{4}
p_0(1-p_0)$, so

\begin{align}
  \var(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}

@@

\begin{align}
  \var(\Delta x_1) &= \var(K_1) + \var(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} \frac{\var(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \\
                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\var(k)}{4} + \frac{1}{2}\right)
\end{align}

(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
Note that if the reproduction process is a neutral multinomial Wright--Fisher,
$\var(k) \approx 2$, and this simplifies to the expected Wright--Fisher
variance. If we were to \emph{define} a variance effective population size by
setting this variance in neutral allele frequency change to the expected
variance under a Wright--Fisher model ($V_\text{WF}$), we'd have

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \\
  V_\text{WF} &= \var(\Delta p_1) \\
  N_e &= \frac{4N}{\var(k) + 2} \\
\end{align}

(c.f. \cite{Wright1938-tv}).

Now, we look at this variance equation behaves with $t > 1$.

\begin{align}
  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
\end{align}

and

\begin{align}
  \label{eq:var_wf}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] + \E_1\left[ {\Delta x_1}^2\right] \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + \mathcal{C}_{1,2} \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + \mathcal{C}_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalesce each generation
and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up due to associations
with heritable fitness backgrounds. Note that the covariance term $\mathcal{C}_{1,2} =
\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] = 0$ under neutral
evolution, and this simplifies to the well-known equation for variance in a
Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\var(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \varepsilon_i$, then
$\var(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
individuals, $V_h > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

@@

\begin{align}
  \var(\Delta x_1) &= \var(S_1) + \var(D_1) + \var(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2}. \\
\end{align}

(c.f. \cite{Santiago1995-hx} equation 11).

To see how the covariance terms accumulate, consider $\var(x_3 - p_0)$. Note
that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
arbitrary, so by symmetry the expected change is zero. Then,

\begin{align}
  \label{eq:var_x}
  \var(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \\
                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2)  \\
                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3) \\
                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2)
\end{align}

The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
allele frequency change from a neutral allele becoming associated with a
fitness background in generation $i$ and that fitness association persisting
until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
are zero since associations in the future cannot affect the past (see p. 1041
of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,

\begin{align}
  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \\
  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
\end{align}

which are chance covariances (across the population, \emph{not} evolutionary
replicates) created by the random sorting of neutral alleles and fitness into
individuals, \emph{each generation}. These covariances are equivalent to the
Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
predict the change in neutral frequency due to heritable fitness (and likewise
with the change due to non-heritable factors).

@@

\begin{align}
  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right)  \\
                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right)  \\
                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'')  \\
                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'')  \\
                               &= S_1' + S_1''
\end{align}

which follows from symmetry, since it does not matter which gamete carrying the
neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The primes now
indicate whether the selected locus is on the same gamete as the neutral
allele (single prime), or the homologous gamete (double prime).

@@

\begin{align}
  S_{2}' &= S_1 (1-\kappa)(1-r) \\
  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \\
  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
\end{align}

Then, the cumulative effect of the fitness associations created in generation
$1$ can be written as,

\begin{align}
  \label{eq:qterm}
  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
\end{align}

or for continuous $t$,

\begin{align}
  \label{eq:q_t}
  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
\end{align}
which as $t \to \infty$, converges to,

@@

\begin{align}
  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \\
  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
\end{align}

With these covariance terms, we now have a full model for $\var(p_t - p_0)$. We
will look at the case for $t=2$. As shown in equation XXX, the second moment of
every term is a function of the variance in neutral allele frequency in that
generation, $x_t(1-x_t)$. This variance does not stay its initial value of
$x_0(1-x_0)$; it decays each generation due to coalesce from the drift and
selection processes modeled by this approach. The rate at which this decay
happens is the effective population size each generation implied by this model. 
We can write,

\begin{align}
    \label{eq:vardecay}
  \var(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \\
                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \\
  \frac{\var(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \\ 
                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
\end{align}

In general, the evolution of the variance in allele frequency change in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

\begin{align}
  \label{eqn:var_xt}
  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^\infty \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{t=1}^{i-1} \left(1-\frac{1}{2 N_{e,t}}\right). \\
\end{align}

Since the probability of coalesce (or, equivalently, identity by descent) is
proportional to the variance in allele frequency change
\parencite{Barton2000-zg}, this encodes the pairwise coalescent rate of the
population through time under selection through the $N_{e,t}$ terms. Others
have found that the genealogies under purifying selection can be characterized
by such a time-dependent effective population size
\parencite{Nicolaisen2013-gv}. In the original derivation of
\textcite{Santiago1995-hx}, they set each $N_{e,t}$ term to a single $N$ and
solve this summation in the limit to infinity giving,

\begin{align}
  \label{eq:var_asymp}
  \frac{\var(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
\end{align}

To relate this to an asymptotic effective population size, note that in an
ideal Wright--Fisher population,

\begin{align}
  \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right]
\end{align}

such that if we look at the change in the variance in allele frequency change
between adjacent generations $t$ and $t-1$, we get an implied $N_{e,t}$

\begin{align}
  \label{eqn:ne_var}
  N_{e,t} = \frac{p_0(1-p_0) - \var(x_{t-1})}{2(\var(x_t) - \var(x_{t-1}))}
\end{align}
%
(cf. p. 1018, \cite{Santiago1995-hx}).

Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
population size,

\begin{align}
  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
  N_e &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
\end{align}

(cf. equation 18, \cite{Santiago1995-hx}). Levels of pairwise diversity,
however, depend on the full distribution of pairwise coalesce times, or the
$N_{e,t}$. While this would require solving the recursion defined by equations
\eqref{eqn:var_xt} and \eqref{eqn:ne_var}, \textcite{Santiago1998-bs} show that
a reasonably good approximation follows from using a single effective
population size for all timepoints, $N_{e,t} = N_e$, but use the time-indexed
$Q_t$. This leads to,

\begin{align}
  \label{eq:ne_t}
  N_{e,t} &\approx \frac{4N}{4 V_h Q_t^2 +V_n+2}
\end{align}

(c.f. equation 15 of \cite{Santiago1995-hx} and p. 1270 of \cite{Santiago2016-mu}).

\subsection{Effective Population Size Under Polygenic Selection}

The results of the previous section assume only a single locus a recombination
fraction $r$ apart from the neutral locus is determining the coalesce rate. The
model of \textcite{Santiago1998-bs} extends this to the case where fitness is
polygenic. As they show in their original paper, this system can accommodate a
variety of different equilibrium selection systems as long as expressions for
$V_h$ and $1-\kappa$ can be worked out. Other authors have extended similar models to
more elaborate bbreeding structures \parencite{Wray1990-zf,Woolliams1993-qo}.

>>>>

\documentclass[11pt]{article}
\RequirePackage{fullpage}
%\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{graphicx}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{wasysym}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{xr}
%\RequirePackage[osf]{mathpazo}
\let\temp\rmdefault
\RequirePackage{mathpazo}
\let\rmdefault\temp

@@

% for cross references
\externaldocument{manuscript}

@@

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\var}{V}
\newcommand{\V}{\text{V}}
\DeclareMathOperator{\cov}{cov}

@@

\section{Quantitative Genetic Linked Selection Model Theory}

@@

The frequency in the initial generation is $x_0 = \nicefrac{1}{N} \sum_{i=1}^N
x_{0,i}$, though to indicate that we treat this as fixed rather than random, we
use $p_0 := x_0$. Then, the allele frequency change is,

\begin{align}
  \Delta x_1 = x_1 - p_0 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (k_{0,i} - 1) + \frac{1}{2N} \sum_{i=1}^N \sum_{j=1}^{k_{0,i}} \delta_{0,i,j} \nonumber \\
  \Delta x_1 &= K_1 + H_1
\end{align}

@@

Now, let us look the variance of $\V(\Delta x_1)$ over evolutionary
replicates. Since the Mendelian segregation and the offspring random process
are independent,

\begin{align}
  \V(\Delta x_1) &= \V(K_1) + \V(H_1).
\end{align}
%
Looking at each term,

\begin{align}
  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V\left(x_{0,i} (k_{0,i} - 1) \right)
\end{align}
%
where we ignore the covariance terms due to the sum, since these are on order
$\nicefrac{1}{N^3}$. In the first generation from an arbitrary starting point,
the neutral alleles assort independently into diploids with respect to their
fitness, so we can simply the variance as

\begin{align}
  \V(K_1) &\approx \frac{1}{4N^2} \sum_{i=1}^N \V(x_{0,i}) \V(k_{0,i} - 1) \nonumber  \\
            &\approx \frac{1}{4N} \V(x_{0,i}) \V(k_{0,i}).
\end{align}

Assuming no correlation between parental gametes (e.g. no inbreeding),
$\V(x_{0,i})$ is the binomial variance in individual allele frequency, or
$p_0(1-p_0)/2$, and $\V(k) := \V(k_{0,i})$ is the offspring variance of an
individual given by the reproduction process. For example, if the reproduction
process is a neutral multinomial Wright--Fisher, $\V(k) \approx 2$. Then, the
variance in allele frequency change due to non-heritable offspring number
variation is,

\begin{align}
  \V(K_1) &\approx \frac{p_0(1-p_0)}{2N} \frac{\V(k)}{4}.
\end{align}

The Mendelian noise variance term can be derived similarly. First, the sum over
each parent's transmitted gametes can be simplified by noting that parents are
exchangeable over evolutionary replicates with respect to their contribution to
this term. In a constant population size, the double summation over parents and
their offspring can be replaced by a summation over offspring, since both sum
$N$ exchangeable terms. Then, note that $\V(\delta_{i,j}) = \nicefrac{1}{4}
p_0(1-p_0)$, so

\begin{align}
  \V(H_1) &= \frac{p_0(1-p_0)}{2N} \frac{1}{2}.
\end{align}

@@

\begin{align}
  \V(\Delta x_1) &= \V(K_1) + \V(H_1) \\
  &\approx \frac{p_0(1-p_0)}{2N} \frac{\V(k)}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \nonumber \\
                   &\approx \frac{p_0(1-p_0)}{2N}\left(\frac{\V(k)}{4} + \frac{1}{2}\right)
\end{align}
%
(c.f. \cite{Santiago1995-hx} equation 2 and \cite{Buffalo2019-qs} equation 30).
Note that if the reproduction process is a neutral multinomial Wright--Fisher,
$\V(k) \approx 2$, and this simplifies to the expected Wright--Fisher
variance. If we were to define a variance \emph{drift-effective} population
size $N_e$ by setting this variance in neutral allele frequency change to the
expected variance under a Wright--Fisher model ($V_\text{WF}$), we'd have

\begin{align}
  V_\text{WF} &:= \frac{p_0(1-p_0)}{2N_e} \nonumber \\
  V_\text{WF} &= \V(\Delta p_1) \nonumber \\
  N_e &= \frac{4N}{\V(k) + 2}
\end{align}
%
(c.f. \cite{Wright1938-tv}). Now, we look at the variance in allele frequency
change across two generations for a neutral allele. The change in neutral
allele frequency, regardless of the selective system, is directionless in
expectation. This is because whichever allele we track is arbitrary by the
definition of neutrality, which imposes a symmetry, $\E(\Delta x) = 0$. Then, the
variance of change in the first generation is,

\begin{align}
  \V(\Delta x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
\end{align}
%
and across two generations,

\begin{align}
  \label{eq:var_wf}
  \V(x_2 - x_0) &= \E\left[(x_2 - x_0)^2\right] \nonumber \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \nonumber \\
          &= \E_1\left[\E_2\left[(\Delta x_2 + \Delta x_1)^2 | x_1\right]\right] \nonumber \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + 2\E_1\left[\E_2[\Delta x_2 \Delta x_1 | x_1]\right] \nonumber \\
          &= \E_1\left[\E_2\left[{\Delta x_2}^2|x_1\right]\right] + \E_1\left[ {\Delta x_1}^2\right] + 2\mathcal{C}_{1,2} \nonumber \\
          &= \frac{p_0(1-p_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{p_0(1-p_0)}{2N} + 2\mathcal{C}_{1,2}
\end{align}

Consequently, the variance of the neutral allele frequency changes each
generation according to the probability of failing to coalescence each
generation and the pairwise covariance terms $\mathcal{C}_{i,j}$ that build up
due to associations with heritable fitness backgrounds. Note that the
covariance term $\mathcal{C}_{1,2} = \E_1\left[\E_2[\Delta x_2 \Delta x_1 |
x_1]\right] = 0$ under neutral evolution, and this simplifies to the well-known
equation for variance in a Wright--Fisher population.

Now, we turn our attention to the case where there is heritable fitness
variation, which leads to non-zero covariance terms $\mathcal{C}_{i,j}$. These
terms emerge when $\V(k)$ has a heritable component of fitness that can be
transmitted along with the neutral allele, thereby affecting the neutral
allele's trajectory in later generations. If we partition the offspring number
into heritable and non-heritable components, $k_i = 2f_i + \Vepsilon_i$, then
$\V(k_i) = 4 V_h + V_n$. When there is heritable fitness variation across
individuals, $V_h > 0$. Because the population size is assumed to be constant,
the mean heritable fitness across individuals is constrained to be $\E_i(f_i) =
1$; this implies that the individual fitness values $f_i$ are relative
fitnesses as is standard in population genetics, in this case to the population
mean. Thus $V_h$ is the squared coefficient of heritable fitness variation
(this is often denoted as $C^2 = \nicefrac{V_A}{\bar{w}^2}$, see
\cite{Crow1958-pc,Charlesworth1987-ab,Houle1992-ur}).

@@

\begin{align}
  \V(\Delta x_1) &= \V(S_1) + \V(D_1) + \V(H_1) \nonumber \\
                   &\approx \frac{p_0(1-p_0)}{2N} V_h + \frac{p_0(1-p_0)}{2N} \frac{V_n}{4} + \frac{p_0(1-p_0)}{2N} \frac{1}{2} \label{supp-eq:varterms} 
\end{align}
%
(c.f. \cite{Santiago1995-hx} equation 11).

To see how the covariance terms accumulate, consider $\V(x_3 - p_0)$. Note
that $\E(S_t) = \E(D_t) = \E(H_3) = 0$, since which neutral allele we track is
arbitrary, so by symmetry the expected change is zero. Then,

\begin{align}
  \label{eq:var_x}
  \V(x_3 - p_0) &= \E\left[ \left(S_1 + D_1 + H_1 + S_2 + D_2 + H_2 + S_3 + D_3 + H_3 \right)^2 \right] \nonumber \\
                  &= \E({S_1}^2) + \E({S_2}^2) + \E({S_3}^2) \nonumber \\
                  &+ \E(S_1 S_2) + \E(S_1 S_3) + \E(S_2 S_3)\nonumber \\
                  &+ \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) + \E({D_3}^2) + \E({H_3}^2).
\end{align}
%
The covariance terms $\E(S_i S_j)$ for $j > i$ represent the expected neutral
allele frequency change from a neutral allele becoming associated with a
fitness background in generation $i$ and that fitness association persisting
until generation $j$. However, the covariances terms $\E(S_j S_i)$ for $j > i$
are zero since associations in the future cannot affect the past (see p. 1041
of \cite{Buffalo2019-qs}). Let us consider $S_1$ and $S_2$. We have,

\begin{align}
  S_1 = \frac{1}{2N} \sum_{i=1}^N x_{0,i}(f_{0,i} - 1) = \cov(x_{0,i}, f_{0,i}) \nonumber \\
  S_2 = \frac{1}{2N} \sum_{i=1}^N x_{1,i}(f_{1,i} - 1) = \cov(x_{1,i}, f_{0,i})
\end{align}
%
which are chance covariances (across the population, \emph{not} evolutionary
replicates) created by the random sorting of neutral alleles and fitness into
individuals, \emph{each generation}. These covariances are equivalent to the
Robertson-Price equation \parencite{Robertson1966-fs,Price1970-si}, which
predict the change in neutral frequency due to heritable fitness (and likewise
with the change due to non-heritable factors).

@@

\begin{align}
  S_1 = \cov(x_{0,i}, f_{0,i}) &= \cov\left(\frac{x_{0,i}' + x_{0,i}''}{2}, f_{0,i}' + f_{0,i}''\right) \nonumber \\
                               &= \frac{1}{2}\left(\cov(x_{0,i}', f_{0,i}' + f_{0,i}'') + \cov(x_{0,i}'', f_{0,i}' + f_{0,i}'')\right) \nonumber  \\
                               &= \cov(x_{0,i}', f_{0,i}' + f_{0,i}'') \nonumber \\
                               &= \cov(x_{0,i}', f_{0,i}') + \cov(x_{0,i}', f_{0,i}'') \nonumber \\
                               &= S_1' + S_1''
\end{align}
%
which follows from symmetry again, since it does not matter which gamete
carrying the neutral allele we track (c.f. \cite{Santiago1998-bs} p. 2107). The
primes now indicate whether the selected locus is on the same gamete as the
neutral allele (single prime), or the homologous gamete (double prime).

@@

\begin{align}
  S_{2}' &= S_1 (1-\kappa)(1-r) \nonumber \\
  S_{3}' &= S_1 (1-\kappa)^2(1-r)^2 \nonumber \\
  S_{4}' &= S_1 (1-\kappa)^3(1-r)^3.
\end{align}
%
Then, the cumulative effect of the fitness associations created in generation
$1$ can be written as,

\begin{align}
  \label{eq:qterm}
  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \nonumber \\
  Q_t &= 1 + \sum_{i=1}^t (1-r)^i(1-\kappa)^i
\end{align}
%
or for continuous $t$, 
%
\begin{align}
  \label{eq:q_t}
  Q_t &= 1+\frac{(1-\kappa) (1-r) \left(1-(1-\kappa)^t (1-r)^t\right)}{\kappa+(1-\kappa)r}
\end{align}
which as $t \to \infty$, converges to,

@@

\begin{align}
  \E(S_1 S_2) &= \E(S_1^2) (1-\kappa)(1-r) = \E(S_1^2) Q_2^2 \nonumber \\
  \E(S_1 S_3) &= \E(S_1^2) (1-\kappa)^2(1-r)^2 = \E(S_1^2) Q_3^2.
\end{align}
%
The $Q_t^2$ terms represent the inflation of variance due to autocovariance in
frequency change across generations due to weak draft.

With these covariance terms, we now have a full model for $\V(p_t - p_0)$. We
will look at the case for $t=2$. As shown in equation \eqref{supp-eq:varterms},
the second moment of every term is a function of the variance in neutral allele
frequency in that generation, $x_t(1-x_t)$. This variance does not stay its
initial value of $x_0(1-x_0)$; it decays each generation due to coalescence
from the drift and selection processes modeled by this approach. The rate at
which this decay happens is the effective population size each generation
implied by this model. We can write,

\begin{align}
    \label{eq:vardecay}
  \V(x_2 - p_0) &= \E({S_1}^2) + \E(S_1 S_2) + \E({S_2}^2) \nonumber \\
                  & + \E({D_1}^2) + \E({H_1}^2) + \E({D_2}^2) + \E({H_2}^2) \nonumber \\
  \frac{\V(x_2 - p_0)}{p_0(1-p_0)} &= \frac{V_h}{2N_{e,1}} + \frac{V_h Q_2^2}{2N_{e,2}}\left(1-\frac{1}{2N_{e,1}}\right)  +  \frac{V_h}{2N_{e,2}}  \left(1-\frac{1}{2N_{e,1}}\right)  \nonumber \\ 
                                     &+ \frac{V_n}{8N_{e,1}} + \frac{1}{4N_{e,1}} + \frac{V_n}{8N_{e,2}}  \left(1 - \frac{1}{2N_{e,1}}\right) 
+ \frac{1}{4N_{e,2}} \left(1 - \frac{1}{2N_{e,1}}\right).
\end{align}
%
In general, the variance in allele frequency change after $t$ generations in a
system with heritable fitness due to a single locus $r$ recombination fraction
away from the neutral site is given by,

\begin{align}
  \label{eqn:var_xt}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\prod_{j=1}^{i-1} \left(1-\frac{1}{2 N_{e,j}}\right).
\end{align}

We will see later that with the correct form for $Q_t^2$, this equation holds
for polygenic selection (see Supplementary Section \ref{supp:ne-polygenic}).
Additionally, we note that since the probability of coalescence (or,
equivalently, identity by descent) is proportional to the variance in allele
frequency change \parencite{Barton2000-zg}, this equation for the variance
encodes the pairwise coalescent rate of the population through time under weak
draft in the $N_{e,t}$ terms. Others have found that the genealogies under
purifying selection can be characterized by such a time-dependent effective
population size \parencite{Nicolaisen2013-gv}.

Now, we consider how to turn this expression for the variance in neutral allele
frequency at time $t$ under weak draft into a \emph{draft-effective} population
size. In an ideal Wright--Fisher population, the variance after $t$ generations
is

\begin{align}
    \label{supp-eq:drift-vt}
    \V(x_t) &= p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_e}\right)^t \right].
\end{align}

To find an analogous expression for the asymptotic \emph{draft-effective}
population size $N_d$, we let $N_e = N_d$ in the above equation, and consider
the change in variance between generations $t$ and $t-1$, $\V(x_t) -
\V(x_{t-1})$. Using Supplementary Equation \eqref{supp-eq:drift-vt}, we have,

\begin{align}
    \V(x_t) - \V(x_{t-1}) = p_0(1-p_0) \left[ \left(1 - \frac{1}{2N_e}\right)^{t-1} - \left(1 - \frac{1}{2N_e}\right)^t  \right]
\end{align}

\textcite{Santiago1995-hx} assume the change in population size per generation
is small enough that the coalescence term each generation is constant, i.e. $(1
- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.

\begin{align}
    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_e} + \frac{V_n}{8N_e} + \frac{1}{4N_e} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
\end{align}

One

Then, imagine the process burns in for $t$ generations and asking what the
equivalent

\begin{align}
    \V(x_t) &= \V(x_{t-1})

  p_0(1-p_0) \left[1 - \left(1 - \frac{1}{2N_d}\right)^t \right].
\end{align}

\begin{align}
  \label{eqn:ne_var}
  N_{d,t} = \frac{p_0(1-p_0) - \V(x_{t-1})}{2(\V(x_t) - \V(x_{t-1}))}
\end{align}
%
(cf. p. 1018, \cite{Santiago1995-hx}).

XXX here

\textcite{Santiago1995-hx} assume the change in population size per generation
is small enough that the coalescence term each generation is constant, i.e. $(1
- \nicefrac{1}{2N_{e,t}}) \approx (1 - \nicefrac{1}{2N_e})$ for each $t$.

\begin{align}
    \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \sum_{i=1}^t \left(\frac{V_h Q_i^2}{2 N_{e,i}} + \frac{V_n}{8N_{e,i}} + \frac{1}{4N_{e,i}} \right)\left(1-\frac{1}{2 N_{e}}\right)^{i-1}.
\end{align}

\begin{align}
  \label{eq:var_asymp}
  \frac{\V(x_t - p_0)}{p_0(1-p_0)} &= \frac{4 V_h Q_\infty^2 +V_n+2}{4 (2 N-1)}.
\end{align}

Note here, to differentiate from the drift-effective population size $N_e$ due
under non-selective causes, we use $N_d$ as the \emph{draft-effective}
population size. Then, if we look at the change in the variance in allele
frequency change between adjacent generations $t$ and $t-1$, we get an implied
$N_{d,t}$

---
Solving this for equation \eqref{eq:var_asymp} gives the asymptotic effective
population size,

\begin{align}
  %N_e &= \frac{2 (2 N-1)}{4 V_h Q_\infty^2 +V_n+2} \\
  N_d &\approx \frac{4N}{4 V_h Q_\infty^2 +V_n+2}
\end{align}
%
(cf. equation 18, \cite{Santiago1995-hx}). Note if $V_h = 0$, this equation
reduces to Wright's (\citeyear{Wright1938-tv}) equation.
